---
title: Chapter 4.4 Summarizing a Probability Distribution
author: Jim Albert and Monika Hu
date: Chapter 4 Discrete Distributions
output: 
  beamer_presentation: default
fontsize: 12pt
---

## Back to Coin-Flipping Example 

- Once we have constructed a probability distribution, we can use this to find probabilities.

- What is the chance that Peter will win at least \$5 in this game?  Winning "at least \$5" includes the possible values $G$ = 6 and $G$ = 10.

-  One finds the probability of interest by adding the probabilities of the individual values.
\begin{align*}
P(G \ge 5) &= P(G = 6 \, \, {\rm or} \, \,  G = 10) \\
&= P(G = 6) + P(G = 10) \\
&= \frac{5+1}{32} = \frac{6}{32}.
\end{align*}

## Another Probability Computation

- What is the probability Peter wins money in this game?  Peter wins money if the gain $G$ is positive and this corresponds to the values $G = 2, 6, 10$.  

- So probability that Peter wins money is

\begin{align*}
P({\rm Peter \, \, wins}) &= P(G > 0) \\
&= P(G = 2) + P(G = 6) + P(G = 10) \\
&= \frac{10 + 5+1}{32} = \frac{1}{2}.
\end{align*}

## Summaries

- It is helpful to compute an "average" of a probability distribution.

- A common measure of "average" is the mean or expected value of $X$, denoted $\mu$ or $E(X)$. The mean is found by 

1. Computing the product of a value of $X$ and the corresponding value of the pmf $f(x) = P(X = x)$ for all values of $X$.
2. Summing the products.

- The formula is:
$$
 \mu = \sum_x x  f(x).
$$

## Example

The computation of the mean for the Peter-Paul game is illustrated below.  

- One sees that the mean of $G$ is $\mu = 0$.

 | \(g\)   | \(P(G = g)\) | \(g \times P(G = g)\) |
| :------ | -----------: | --------------------: |
| \(-10\) |         1/32 |            \(-10/32\) |
| \(-6\)  |         5/32 |            \(-30/32\) |
| \(-2\)  |        10/32 |            \(-20/32\) |
| 2       |        10/32 |                 20/32 |
| 6       |         5/32 |                 30/32 |
| 10      |         1/32 |                 10/32 |
| SUM     |            1 |                     0 |


## Interpretation

- How does one interpret a mean value of 0?  

- Note that $G = 0$ is not a possible outcome of the game.  

- But if Peter and Paul play this game a large number of times, then the value $\mu$= 0 represents (approximately) the mean winnings of Peter in all of these games.


## Simulating the Peter-Paul Game (continued)

- The functions ```sample()```\index{sample()} and ```replicate()```\index{replicate()} were earlier illustrated to simulate this game 1000 times in R. 

- Peter's winnings in the different games are stored in the vector ```G```.  Here is a display of Peter's winnings in the first 100 games:

```{r, echo = FALSE}
one_play <- function(){
  flips <- sample(c("H", "T"), 
                  size = 5,
                  replace = TRUE)
  2 * sum(flips == "H") -
       2 * sum(flips == "T")
}
```

```{r, echo = FALSE}
G <- replicate(1000, one_play())
```

```{r}
G[1:100]
```

## Compute Sample Mean

- One approximates the mean winning $\mu$ by finding the sample mean $\bar G$ of the winning values in the 1000  games.

```{r}
mean(G)
```

- This value is approximately equal to the mean of $G$, $\mu$= 0.  

- If Peter was able to play this game for a much larger number of games,  his average winning  would be very close to $\mu = 0$.


