---
title: Chapter 11.9 A Conditional Means Prior
author: Jim Albert and Monika Hu
date: Chapter 11 Simple Linear Regression
output: 
    beamer_presentation: default
    logo: ProbBayes_cover.jpg
fontsize: 12pt
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
library(runjags)
library(coda)
library(tidyverse)
library(ggridges)
crcblue <- "#2905a1"
```


## Recap

- Two methods for constructing  a prior on the parameters of a regression model so far:
    - a weakly informative prior (on the original data)
    - an informative prior (on the standardized data)
  
- A new approach
    - on the original data
    - instead of a prior on $(\beta_0, \beta_1, \sigma)$
    - prior about  the expected response value conditional on specific values of the predictor variable


## Learning about a gas bill from the outside temperature

- Can one accurately predict one's monthly natural gas bill from the outside temperature?

- Outcome / response variable: montly natural gas bill

- Explanatory / predictor variable: outside temperature

\pause 

\begin{equation}
Y_i \mid \beta_0, \beta_1, \sigma \sim {\rm Normal}(\beta_0 + \beta_1 x_i, \sigma)
\end{equation}

- $x_i$ and $y_i$ are respectively the average temperature (degrees in Fahrenheit) and the bill amount (in dollars) in month $i$

- $(\beta_0, \beta_1, \sigma)$ are the unknown regression parameters

## Learning about a gas bill from the outside temperature cont'd

```{r size = "footnotesize", echo = FALSE, fig.height = 3, fig.width = 4.5, warning = FALSE, message = FALSE}
gas_prices <- read_csv("gas2017.csv")

ggplot(gas_prices, aes(Temp, Bill)) +
  geom_point(size = 1, color = crcblue) +
  xlab("Average Monthly Temperature (degrees F)") +
  ylab("Monthly Bill (dollars)") +
  theme(text = element_text(size=10))
```

## A conditional means prior

- Assume independence of parameters, as before
\begin{equation*}
\pi(\beta_0, \beta_1, \sigma) = \pi(\beta_0, \beta_1) \pi(\sigma)
\end{equation*}

- Think about formulating prior opinion about the mean values
\begin{equation}
\mu_i^* = \beta_0 + \beta_1 x_i^*
\end{equation}

- For two specified values of the predictor $x_1^*$ and $x_2^*$

## A conditional means prior cont'd

Two steps:

1. For the first predictor value $x_1^*$ construct a normal prior for the mean value $\mu_1^*$
    - let the mean and standard deviation values of this prior be denoted by $m_1$ and $s_1$, respectively
2. Similarly, for the second predictor value $x_2^*$  construct a normal prior for the mean value $\mu_2^*$ 
    - let the mean and standard deviation values of this prior be denoted by $m_2$ and $s_2$

## A conditional means prior cont'd

- Under independence, the joint prior for the vector $(\mu_1^*, \mu_2^*)$ has the form
\begin{equation*}
\pi(\mu_1^*, \mu_2^*) = \pi(\mu_1^*) \pi(\mu_2^*)
\end{equation*}

\pause

- This prior on the two conditional means implies a bivariate normal prior on the regression parameters
\begin{equation}
\beta_1 = \frac{\mu_2^* - \mu_1^*}{x_2 - x_1}
\end{equation}
\begin{equation}
\beta_0 = \mu_1^* - x_1 \left(\frac{\mu_2^* - \mu_1^*}{x_2 - x_1}\right)
\end{equation}

- The slope $\beta_0$ and $\beta_1$ are linear functions of the two conditional means $\mu_1^*$ and $\mu_2^*$, implying a bivariate normal distribution for $\beta_0$ and $\beta_1$

## Regression analysis of the gas bill example

- Consider two different temperature values, say 40 degrees and 60 degrees

- If $x = 40$, the mean bill $\mu_1^* = \beta_0 + \beta_1 (40)$ 
    - a normal prior with mean \$100 and standard deviation \$20
    - we believe the average gas bill will be relatively high during a cold month averaging 40 degrees

- If $x = 60$, the  mean bill $\mu_2^* = \beta_0 + \beta_1 (100)$
    - a normal with mean \$50 and standard deviation \$15
    - we believe the gas cost will average \$50 lower than in the first scenario

\pause

- Under independence
\begin{equation}
\pi(\mu_1^*, \mu_2^*) = \phi(\mu_1^*, 100, 20) \phi(\mu_2^*, 50, 15),
\end{equation}
where $\phi(y, \mu, \sigma)$ denotes the normal density with mean $\mu$ and standard deviation $\sigma$

## Regression analysis of the gas bill example cont'd

- The prior on the two means is an indirect way of assessing a prior on the regression parameters $\beta_0$ and $\beta_1$

- We can  simulate pairs $(\beta_0, \beta_1)$ from the prior distribution by simulating values of the means $\mu_1^*$ and $\mu_2^*$ from independent normal distributions and applying Equation (3) and Equation (4)

```{r size = "footnotesize"}
modelString = "
model{
beta1 <- (mu2 - mu1) / (x2 - x1)
beta0 <- mu1 - x1 * (mu2 - mu1) / (x2 - x1)
mu1 ~ dnorm(m1, s1)
mu2 ~ dnorm(m2, s2)
}"
```

## Regression analysis of the gas bill example cont'd

```{r size = "footnotesize", echo = FALSE, fig.height = 2.5, fig.width = 4, warning = FALSE, message = FALSE, results = 'hide'}
the_data <- list("x1" = 40,
                 "x2" = 60,
                 "m1" = 100,
                 "m2" = 50,
                 "s1" = 1 / 20 ^ 2,
                 "s2" = 1 / 15 ^ 2)

initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

prior <- run.jags(modelString,
                       n.chains = 1,
                       data = the_data,
                       monitor = c("beta0", "beta1"),
                       adapt = 1000,
                       burnin = 5000,
                       sample = 5000,
                       inits = initsfunction)

prior %>% as.mcmc() %>% as.data.frame() -> prior_df


ggplot(prior_df, aes(beta0, beta1)) +
  geom_point(color = crcblue, alpha = 0.2) +
  increasefont(Size = 10) +
  xlab("beta0") +
  ylab("beta1")

```

- 1000 simulated draws of $(\beta_0, \beta_1)$ 
- The implied prior on the regression coefficients indicates that $\beta_0$ and $\beta_1$ are strongly negatively correlated

## Regression analysis of the gas bill example cont'd

- Don't forget to assign the precision parameter $\phi = 1 / \sigma^2$ a gamma prior with parameters $a$ and $b$
\begin{equation*}
\pi(\beta_0, \beta_1, \sigma)  = \pi_{CM}(\beta_0, \beta_1) \pi(\sigma),
\end{equation*}
where $\pi_{CM}$ is the conditional means prior

\pause

- Use JAGS for posterior inference

- Compare inferences using conditional means and weakly informative priors
