---
title: Chapter 11.6 Inference through MCMC
author: Jim Albert and Monika Hu
date: Chapter 11 Simple Linear Regression
output: 
    beamer_presentation: default
    logo: ProbBayes_cover.jpg
fontsize: 12pt
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
library(runjags)
library(coda)
library(tidyverse)
crcblue <- "#2905a1"
```

## Overview

- As before, we will use JAGS to draw MCMC samples (Chapter 9 and Chapter 10)

- Straightforward to transpose the statement of the Bayesian model (sampling density and prior) directly to the JAGS model script

```{r message = FALSE, warning = FALSE, size = "footnotesize"}
PriceAreaData <- read_csv("house_prices.csv")
PriceAreaData$newsize <- PriceAreaData$size / 1000
```

## JAGS step 1: describe the model by a script


```{r size = "footnotesize"}
modelString <-"
model {
## sampling
for (i in 1:N){
   y[i] ~ dnorm(beta0 + beta1*x[i], invsigma2)
}
## priors
beta0 ~ dnorm(mu0, g0)
beta1 ~ dnorm(mu1, g1)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}"
```


## JAGS step 2: define the data and prior parameters


```{r size = "footnotesize"}
y <- PriceAreaData$price  
x <- PriceAreaData$newsize   
N <- length(y)  
the_data <- list("y" = y, "x" = x, "N" = N,
                 "mu0" = 0, "g0" = 0.0001,
                 "mu1" = 0, "g1" = 0.0001,
                 "a" = 1, "b" = 1)
```

## JAGS step 3: generate samples from the posterior distribution

```{r size = "footnotesize", message = FALSE, warning = FALSE, results = 'hide'}
posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", 
                                  "beta1", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000)
```

## MCMC diagnostics and summarization

```{r, fig.height = 3, fig.width = 4.5}
plot(posterior, vars = "beta0")
```

## MCMC diagnostics and summarization cont'd

- A scatterplot of the simulated draws of the regression parameters $\beta_0$ and $\beta_1$ (a strong negative correlation)

```{r, fig.height = 3, fig.width = 4.5, echo = FALSE}
post <- as.mcmc(posterior)
ggplot(as.data.frame(post), aes(beta0, beta1)) +
  geom_point(color = crcblue, alpha = 0.2) +
  increasefont(Size = 10) + xlab("beta0") + ylab("beta1")
```


## MCMC diagnostics and summarization cont'd

```{r size = "footnotsize", message = FALSE, warning = FALSE, eval = FALSE}
print(posterior, digits = 3)
```

```
      Lower95 Median Upper95  Mean   SD Mode MCerr 
beta0    -122  -46.2    31.4 -45.7 37.6   --  2.98     
beta1    78.7    117     159   117   20   --  1.65     
sigma    33.2     45    59.3  45.7 6.93   -- 0.157
```

\pause

- Interpretations
    - Intercept $\beta_0$
    - Slope $\beta_1$
    - Standard deviation $\sigma$
