---
title: Chapter 11.2 Example - Prices and Areas of House Sales
author: Jim Albert and Monika Hu
date: Chapter 11 Simple Linear Regression
output: 
    beamer_presentation: default
    logo: ProbBayes_cover.jpg
fontsize: 12pt
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
library(runjags)
library(coda)
library(tidyverse)
crcblue <- "#2905a1"
```

## The dataset

- Zillow: an online real estate database company that collects information on 110 million homes across the United States

- Our sample: a random sample of 24 houses for sale in the Findlay, Ohio area during October 2018

- For each house, the dataset contains the selling price (in \$1000) and size (in 1000 square feet)

## The dataset cont'd

| Index | Price ($1000) | Size (1000 sq feet) |
| :---: | :-----------: | :-----------------: |
|   1   |      167      |        1.625        |
|   2   |      236      |        1.980        |
|   3   |      355      |        2.758        |
|   4   |      148      |        1.341        |
|   5   |      93       |        1.465        |

## The dataset cont'd

```{r, fig.height = 3, fig.width = 4.5, message = FALSE, warning = FALSE, echo = FALSE}
PriceAreaData <- read_csv("house_prices.csv")

PriceAreaData$newsize <- PriceAreaData$size / 1000

ggplot(PriceAreaData, aes(x = newsize, y = price)) +
  geom_point(size=1, color = crcblue) +
  labs(x = "Size (1000 sq feet)", y = "Price ($1,000)") +
  increasefont(Size = 10)
```

## Some questions

- Suppose one is interested in predicting a house's selling price  from its house size

- The scatter plot shows a positive relationship between the size and the price of a house sale, suggesting that the house sale price increases as the house size increases

\pause

- Can one quantify this relationship through a Bayesian linear regression model? 

- Is there sufficient evidence that there is a positive association among the population of all homes?  

- Can one predict the sale price of a home given its size?


