---
title: Chapter 3.1 Introduction
author: Jim Albert and Monika Hu
date: Chapter 3 Conditional Probability
output: 
  beamer_presentation: default
fontsize: 12pt
---

## The Three Card Problem

- Suppose one has three cards -- one card is blue on both sides, one card is pink on both sides, and one card is blue on one side and pink on the other side.  

- Suppose one chooses a card and place it down showing "blue".  What is the chance that the other side is also blue?

- This is a famous conditional probability problem.
One is given certain information -- here the information is that one side of the card is blue -- and one wishes to determine the probability that the other side is blue.

## A simulation experiment

- One can obtain the correct answer by simulating this experiment many times.  

- Think of this experiment as first choosing a card, and then choosing a side from the card.  

- There are three possible cards, which we call "Blue", "Pink" and "mixed".  For the blue card, there are two blue sides; for the pink card, there are two pink sides, and the "mixed" card has a blue side and a pink side.  

## A R simulation

- Define a data frame ```df``` with two variables ```Card``` and ```Side```.

- ```sample()``` function randomly chooses a card and  a side by choosing a random row from the data frame.  Repeat the experiment 1000 times.

```{r, echo = FALSE}
set.seed(123)
```
```{r}
df <- data.frame(Card = c("Blue", "Blue",
                          "Pink", "Pink",
                          "Mixed", "Mixed"),
                 Side = c("Blue", "Blue",
                          "Pink", "Pink",
                          "Blue", "Pink"))
cards <- df[sample(6, size = 1000, replace = TRUE), ]
```

## Table of outcomes

```{r}
table(cards$Card, cards$Side)
```

## What did we learn?

- One observed "side is blue" and one are interested in the probability of the event "card is blue". 

- In this experiment, the blue side was observed 346 + 162 = 508 times.

- Of these blue sides, the card was blue 346 times.  
- So the probability the other side is blue is approximated by 346 / 508  which is close to the exact probability of 2/3.  

## Selecting slips of paper

- Suppose one has a box that has 6 slips of paper labeled with the numbers 2, 4, 6, 8, 10, and 12.  

- One selects two slips at random without replacement from the box.

- Here are the  $_6 C_2 = 15$ possible outcomes

$S$ = \{(2, 4), (2, 6), (2, 8), (2, 10), (2, 12), (4, 6), (4, 8), (4, 10), (4, 12)
(6, 8), (6, 10), (6, 12), (8, 10), (8, 12), (10, 12)\}.

## Finding a probability

- Suppose one are interested in the probability the sum of the numbers on the two slips is 14 or higher.

- Assume that the 15 outcomes listed above are equally likely.

- One sees there are 9 outcomes where the sum is 14 or higher and so 
$$
Prob({\rm sum} \, 14 \, {\rm or} \, {\rm higher}) = \frac{9}{15}.
$$

## New information

- Suppose one is given some new information  --  both of the numbers on the slips are single digits.

- Given this information, one now has only six possible outcomes.  

- This new sample space is called the reduced sample space based on the new information.

$S$ = \{(2, 4), (2, 6), (2, 8), (4, 6), (4, 8), (6, 8)\}

## A conditional probability

- One finds the probability $Prob$(sum is 14 or higher) given that both of the slip numbers are single digits.   

- Since there is only one way of obtaining a sum of 14 or higher in our new sample space, one sees
$$
Prob({\rm sum} \, 14 \, {\rm or} \, {\rm higher}) = \frac{1}{6}.
$$

- Notation:  we write

$Prob$(sum is 14 or higher  |  both numbers are single digits).


- Initially, the probability of 14 and higher was  high (9/15), but given the new information, the probability dropped to 1/6.  
