---
title: Chapter 8.4 Continuous Priors
author: Jim Albert and Monika Hu
date: Chapter 8 Modeling Measurement and Count Data
output: 
  beamer_presentation: default
fontsize: 12pt
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
library(tidyverse)
crcblue <- "#2905a1"
```

## The Normal prior

- One is interested in learning about the time-to-serve for the tennis player Roger Federer.  

- His serving times are  Normally distributed with unknown mean $\mu$ and known standard deviation $\sigma = 4$.  

- One standard approach for representing one's belief about a Normal mean\index{Normal!mean} is based on a Normal prior density with mean $\mu_0$ and standard deviation $\sigma_0$, that is 
\begin{eqnarray*}
\mu \sim {\rm{Normal}}(\mu_0, \sigma_0).
\end{eqnarray*}  

- The value $\mu_0$ represents one's "best guess" at the mean time-to-serve $\mu$ and $\sigma_0$ indicates how sure one thinks about the guess.

## Two priors

- Consider the opinion of Joe who has strong prior information about the mean.  

- His best guess at Federer's mean time-to-serve is 18 seconds so he lets $\mu_0 = 18$.  

- He is very sure of this guess and so he chooses $\sigma_0$ to be the relatively small value of $0.4$.

- In contrast,  Kate also thinks that Federer's mean time-to-serve is 18 seconds, but does not have a strong belief in this guess and chooses the large value $2$ of the standard deviation $\sigma_0$.  

## Plot of two Normal priors for the mean time-to-serve $\mu$.

```{r, echo = FALSE, warning = FALSE, fig.height = 4, message = FALSE, fig.cap = "Two priors for the Normal mean $\\mu$."}
par1 <- c(18, 0.4)
par2 <- c(18, 2)

ggplot(data.frame(x=c(10, 26)), aes(x)) +
  stat_function(fun=dnorm, geom="line",
                aes(linetype="dashed"),
                size=1.5, color = crcblue,
                args=list(mean=par1[1],
                          sd=par1[2])) +
  stat_function(fun=dnorm, geom="line",
                aes(linetype="solid"),
                size=1.5, color = crcblue,
                args=list(mean=par2[1],
                          sd=par2[2])) +
  xlab(expression(mu)) + ylab("Density") +
  theme(legend.position = "none") +
  annotate(geom = "text", x = 22, y = 0.7,
           size = 8, label = "N(18, 0.4)") +
  annotate(geom = "text", x = 23, y = 0.15,
           size = 8, label = "N(18, 2)") +
  increasefont()
```


## Choosing a informative Normal prior {#Normal:Continuous:choosing}


- Want to choose a Normal prior for $\mu$ that reflects prior beliefs about the location of this parameter  

- One indirect strategy for choosing for selecting values $\mu_0$ and $\sigma_0$ is based on the specification of quantiles\index{quantile}.  

- On the basis of one's prior beliefs, one specifies two quantiles of the Normal density\index{Normal!density}.  

- Then the Normal parameters are found by matching these two quantiles to a particular Normal curve\index{Normal!curve}.

## Construction of a  prior

- Suppose one specifies the 0.5 quantile to be 18 seconds --- this means that $\mu$ is equally likely to be smaller or larger than 18 seconds.  

- Next, one decides that the 0.9 quantile is 20 seconds.  This means that one's probability that $\mu$ is smaller than 20 seconds is 90\%.  

- Given values of these two quantiles, the unique Normal curve is found that matches this information.

## Normal prior

- The Normal curve with mean $\mu_0 = 18$ and $\sigma_0 = 1.56$, displayed in Figure 8.5, matches the prior information stated by the two quantiles.  

```{r, echo = FALSE, warning = FALSE, fig.height = 4, message = FALSE, fig.cap = "A person's Normal prior for Federer's mean time-to-serve $\\mu$."}
normal_draw(c(18, 1.56), Color = crcblue) +
  xlab(expression(mu)) +
  ggtitle("") +
  increasefont()
```

## This prior is just an approximation to one's beliefs about $\mu$

- Recommend in practice that one perform several checks to see if this Normal prior makes sense.  

- For example, one finds the 0.25 quantile of our prior using the ```qnorm()```\index{qnorm()} function.

```{r, warning = FALSE, message = FALSE}
qnorm(0.25, 18, 1.56)
```

- The prior probability that $\mu$ is smaller than 16.95 is 25\%. 

- If this does not seem reasonable, make adjustments the values of the Normal mean and standard deviation until a reasonable Normal prior is found.

## Weekly informative prior 

- What would a user do in the situation where little is known about the location on $\mu$?  

- If one is really unsure about any guess at $\mu$, then one assigns the standard deviation $\sigma_0$  a large value.  

- Then the choice of the prior mean will not matter, so we suggest using a Normal(0, $\sigma_0$) with a large value for $\sigma_0$.

- This prior indicates that $\mu$ may plausibly range over a large interval and represents weakly informative prior belief about the parameter.

- The posterior inference for $\mu$ will largely be driven by the data.  



