---
title: Chapter 7.3 Continuous Priors
author: Jim Albert and Monika Hu
date: Chapter 7 Learning About a Binomial Probability
output: 
    beamer_presentation: default
    logo: ProbBayes_cover.jpg
fontsize: 12pt
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
crcblue <- "#2905a1"
```

## Introduction

- $p$ is continuous on [0, 1]

- One possible choice of a continuous prior: the continuous uniform distribution
    - expresses the opinion that $p$ is equally likely to take any value between 0 and 1

- The probability density function of the continuous uniform on the interval $(a, b)$ is
    - $\pi(p) = \frac{1}{b - a} \text{for }a \le p \le b$
    - $\pi(p) = 0 \text{for }p < a \,\, \text{or } p > b$
    
## Beta distribution

- A beta distribution, denoted by Beta($a, b)$, represents probabilities for a random variable falling between 0 and 1

- The beta distribution has two shape parameters\index{shape parameter}, $a$ and $b$, with probability density function given by
$$
\pi(p) = \frac{1}{B(a, b)} p^{a - 1} (1 - p)^{b - 1}, \, \, 0 \le p \le 1,
$$
    - $B(a, b)$ is the beta function: $B(a, b) = \frac{\Gamma(a)\Gamma(b)}{\Gamma(a+b)}$
    - mean $E[p] = \frac{a}{a+b}$
    - variance $V(p) = \frac{ab}{(a+b)^2(a+b+1)}$
    
- $\textrm{Uniform}(0, 1) = \textrm{Beta}(1, 1)$

## Beta distribution cont'd

- Density curves of beta distributions for several choices of the shape parameters

```{r, echo = FALSE, warning = FALSE, message = FALSE}
betapars <- matrix(c(0.5, 0.5,
                     0.5, 1,
                     0.5, 2,
                     1, 0.5,
                     1, 1,
                     1, 2,
                     4, 0.5,
                     4, 1,
                     4, 2),
                     9, 2, byrow = TRUE)
p <- seq(.001, .999, length.out = 100)
BETA <- NULL
for (j in 1:9){
  df <- data.frame(p = p, Density = dbeta(p,
                    betapars[j, 1], betapars[j, 2]))
  df$Type <-  paste("Beta(", betapars[j, 1],
                    ",", betapars[j, 2], ")",
                    sep = "")
  BETA <- rbind(BETA, df)
}
ggplot(BETA, aes(p, Density)) +
  geom_line(color = crcblue, size = 1.5) +
  facet_wrap(~ Type, scale = "free") +
  increasefont() +
  theme(axis.text.y=element_blank()) +
  scale_x_continuous(breaks=seq(0, 1, 0.5))
```

## R for beta distribution

- ```dbeta()```\index{dbeta()}: the probability density function for a Beta(a, b) which takes a value of the random variable as its input and outputs the probability density function at that value

```{r, echo = TRUE, warning = FALSE, message = FALSE}
dbeta(c(0.5, 0.8, 1.2), 1, 1)
```

## R for beta distribution cont'd

- ```pbeta()```: the distribution function of a Beta(a, b) random variable,
which takes a value x and gives the value of the random variable at
that value, $F(x)$

```{r, echo = TRUE, warning = FALSE, message = FALSE}
pbeta(c(0.5, 0.8), 1, 1)
```

One calculates the probability of $p$ between two values by taking the difference of the cdf at the two values

```{r, echo = TRUE, warning = FALSE, message = FALSE}
pbeta(0.8, 1, 1) - pbeta(0.5, 1, 1)
```

## R for beta distribution cont'd

- ```qbeta()```\index{qbeta()}: the quantile function of a Beta(a, b), which inputs a probability value $p$ and outputs the value of $x$ such that $F(x) = p$

```{r, echo = TRUE, warning = FALSE, message = FALSE}
qbeta(c(0.5, 0.8), 1, 1)
```

- ```rbeta()```\index{rbeta()}: the random number generator for Beta(a, b), which inputs the size of a random sample and gives a vector of the simulated random variates


```{r, echo = TRUE, warning = FALSE, message = FALSE}
rbeta(3, 1, 1)
```

## ProbBayes for beta distribution

- Suppose one has a Beta(7, 10) curve and we want to find the chance that $p$ is between 0.4 and 0.8 

```{r, echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3, fig.width = 4}
beta_area(0.4, 0.8, c(7, 10), Color = crcblue)
```


## ProbBayes for beta distribution cont'd

- ```beta_quantile()``` automatically produces a plot with the shaded probability area

```{r, echo = TRUE, warning = FALSE, message = FALSE, fig.height = 3, fig.width = 4}
beta_quantile(0.5, c(7, 10), Color = crcblue)
```


## Choosing a beta density as prior 

- We cover two methods

- The first method is to consider:
    - the shape parameter $a$ as the prior count of "successes"
    - the shape parameter $b$ as the prior count of "failures"
    
- Subsequently, the values:
    - $a + b$ represents the **prior sample size**
    - $n$ represents the **data sample size**
    
## Examples: Beta(4, 4) and Beta(2, 9)

```{r, echo = FALSE, warning = FALSE, message = FALSE}
Beta44samples <- rbeta(1000, 4, 4)
Beta29samples <- rbeta(1000, 2, 9)
df1 <- data.frame(P = Beta44samples, Type = "Beta(4, 4)")
df2 <- data.frame(P = Beta29samples, Type = "Beta(2, 9)")
df <- rbind(df1, df2)
ggplot(df, aes(P)) +
  geom_histogram(fill = crcblue, color = "white",
                 bins = 15) +
  facet_wrap(~ Type, ncol = 1) + increasefont()
```

## Examples: Beta(4, 4) and Beta(2, 9) cont'd

- To further check the quantiles of the prior, use the ```quantile()``` function

```{r, echo = TRUE, warning = FALSE, message = FALSE}
Beta44samples <- rbeta(1000, 4, 4)
quantile(Beta44samples, c(0.25, 0.75))
```

- Discussion: what are the differences and similarities between Beta(4, 4) and Beta(40, 40)?

## Choosing a beta density as prior 

- The second indirect method is by specification of quantiles of the distribution
    - determine the shape parameters $a$ and $b$ by first specifying two quantiles of the beta density curve
    - find the beta density curve that matches these quantiles
    
## Example: specifying the 0.5 and 0.9 quantiles

- The restaurant owner thinks of a value $p_{50}$ such that the proportion $p$ is equally likely to be smaller or larger than $p_{50}$: $p_{50}$ = 0.55

The owner thinks of a value $p_{90}$ that he is pretty sure (with probability 0.90) that the proportion $p$ is smaller than $p_{90}$: $p_{90}$ = 0.80


```{r, echo = TRUE, warning = FALSE, message = FALSE}
beta.select(list(x = 0.55, p = 0.5),
            list(x = 0.80, p = 0.9))
```

## Example: specifying the 0.5 and 0.9 quantiles cont'd

- Check the chosen beta prior
```{r, echo = FALSE, warning = FALSE, message = FALSE}
beta_interval(0.5, c(3.06, 2.56), Color = crcblue) +
  increasefont()
```
