---
title: Chapter 7.1 Introduction - Thinking About A Proportion Subjectively
author: Jim Albert and Monika Hu
date: Chapter 7 Learning About a Binomial Probability
output: 
    beamer_presentation: default
    logo: ProbBayes_cover.jpg
fontsize: 12pt
---


```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(ProbBayes)
library(dplyr)
crcblue <- "#2905a1"
```

## Example: 60 balls in a box

- White balls and red balls

- Do not know the proportions of color balls

- Goal: learn the proportion, $p$, of red balls


## Review: three views of a probability

- The classical view: one needs to write down the sample space where each outcome is equally likely

- The frequency view: one needs to repeat the random experiments many times under identical conditions

- The subjective view: one needs to express one's opinion about the likelihood of a one-time event

- Which one works, why and how?

## Possible values of $p$

- Proportion can take any possible value between 0 and 1

- Example 1: one thinks that $p$ is 0.5
    - i.e. the probability of $p$ = 0.5 is 1
    - could be too strong

- Example 2: $p$ can take 10 different values, denoted by set $A$:
$$A = \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0\}$$

## Probability associated with each possible value of $p$

Though $p$ can take the same 10 multiple values in both scenarios, we assign different probabilities to each possible value.

- Scenario 1:
$$f_1(A) = (0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1)$$

- Scenario 2:
$$f_2(A) = (0.05, 0.05, 0.05, 0.175, 0.175, 0.175, 0.175, 0.05, 0.05, 0.05)$$

## Comparing $f_1(A)$ and $f_2(A)$


```{r, echo = FALSE, warning = FALSE, message = FALSE}
Prior1 <- data.frame(p = seq(0.1, 1, length.out = 10),
                     Probability = rep(0.1, 10),
                     Type = "Prior 1")
Prior2 <- data.frame(p = seq(0.1, 1, length.out = 10),
                     Probability = c(0.05, 0.05, 0.05,
    0.175, 0.175, 0.175, 0.175, 0.05, 0.05, 0.05),
                     Type = "Prior 2")
bothpriors <- rbind(Prior1, Prior2)

ggplot(bothpriors, aes(p, Probability, color = Type)) +
  geom_segment(aes(xend = p, yend = 0), size = 8,
               lineend = "butt",
               color = crcblue) +
  facet_wrap(~Type, ncol = 1) +
  increasefont()
```

## Comparing $f_1(A)$ and $f_2(A)$ cont'd

- The probability assignment in $f_1(A)$ is called a discrete uniform distribution
    - each possible value of the proportion $p$ is equally likely
    - each value gets assigned a probability of $1/10 = 0.1$


- The probability assignment in $f_2(A)$ is also discrete, however, not a Uniform distribution
    - the probabilities of the first three values and last three values are each $1/3.5$ of that of the middle four values
    - the shape of the bins reflects the opinion that the middle values of $p$ are 3.5 times as likely as the extreme values of $p$

## Review: three probability axioms

- Both sets of probabilities follow the three probability axioms in Chapter 1.

- Within each set:
    - each probability is nonnegative;
    - the sum of the probabilities is 1;
    - the probability of mutually exclusive values is the sum of probability of each value


## Bayesian inference

- Step 1: **Prior**: express an opinion about the location of the proportion $p$ before sampling

- Step 2: **Data/Likelihood**: take the sample and record the observed proportion of red balls

- Step 3: **Posterior**: use Bayes' rule to sharpen and update the previous opinion about $p$ given the information from the sample


